---
title: "Temperature and dates"
author: "Christian Escoto"
date: "2025-06-11"
output: pdf_document
---

## Packages
```{r libraries, include=FALSE}
library(dplyr)
library(Matrix)
library(xgboost)
library(caret)
library(purrr)
library(ggplot2)
library(reshape2)
library(viridis)
library(tidyr)
library(ggExtra)
library(openmeteo)

```


## Data preparation
We have separatly the data 

```{r temperature}

termStP22<-read.csv("Temp St.Peter 2022.csv")
termStP23<-read.csv("Temp St.Peter 2023.csv")

termUrb22<-read.csv("Temp Urbana 2022.csv")
termUrb23<-read.csv("Temp Urbana 2023.csv")

termNeo22<-read.csv("Temp Neoga 2022.csv")
termNeo23<-read.csv("Temp Neoga 2023.csv")
```

Reshape the data for the analysis
```{r reshape}
# Reshape the data
termStP22 <- termStP22 %>%
  pivot_wider(names_from = y, values_from = x)

termStP23 <- termStP23 %>%
  pivot_wider(names_from = y, values_from = x)

termUrb22 <- termUrb22 %>%
  pivot_wider(names_from = y, values_from = x)

termUrb23 <- termUrb23 %>%
  pivot_wider(names_from = y, values_from = x)

termNeo22 <- termNeo22 %>%
  pivot_wider(names_from = y, values_from = x)

termNeo23 <- termNeo23 %>%
  pivot_wider(names_from = y, values_from = x)

```

Define the function to calculate GDD 
```{r Define the GDD function}
gdd_calculator <- function(loc, plntDate, dates) {
  library(purrr)
  
  gdd_calculator_helper <- function(date) {
    temps <- as.data.frame(weather_history(
      location = loc,
      start = plntDate,
      end = date,
      daily = c("temperature_2m_max", "temperature_2m_min")
    ))
    vec <- (temps[, 2] + temps[, 3]) / 2
    vec[which(vec < 0)] <- 0
    a_gdd <- sum(vec)
    return(a_gdd)
  }
  
  return(map_dbl(dates, gdd_calculator_helper))
}
```


Function to calculate temperature according with dates and coordinates
```{r}
thermTimeprofile <- function(loc = c(39.31948, -88.45283), plntDate = "2021-10-22", date = "2022-08-14") {
  # Retrieve weather data
  temps <- as.data.frame(weather_history(
    location = loc,
    start = plntDate,
    end = date,
    hourly = 'temperature_2m'
  ))
  
  # Clean data
  temps <- na.omit(temps)
  temps$date <- as.Date(substr(as.character(temps[, 1]), 1, 10))
  temps$time <- trimws(gsub("^\\s*|\\s*$", "", as.character(temps[, 1])))
  
  # Combine date and time
  temps <- temps %>%
    mutate(
      hour = if_else(grepl(':', time), time, paste(time, "00:00:00")),
      julian = yday(date),
      temps = as.numeric(temps[, 2])
    )
  
  # Adjust Julian day for year change
  julian_one_exists <- any(temps$julian == 1)
  
  if (julian_one_exists) {
    temps[1:min(which(temps$julian == 1)) - 1, 'julian'] <- temps[1:min(which(temps$julian == 1)) - 1, 'julian'] - 365
  } else {
    warning("Julian day 1 not found in the data. Adjustments for year change will not be made.")
  }
  
  # Temperature thresholds function
  threshold_temps <- function(temp, lower, upper) {
    temp[temp < lower] <- 0
    temp[temp > upper] <- upper
    return(temp)
  }
  
  temps <- temps %>%
    mutate(
      temps0.26 = threshold_temps(temps, 0, 26),
      temps5.26 = threshold_temps(temps, 5, 26)
    )
  
  # Daily averages
  avgTbl <- temps %>%
    group_by(julian, date) %>%
    summarise(
      temps_mean = mean(temps, na.rm = TRUE),
      temps0.26_mean = mean(temps0.26, na.rm = TRUE),
      temps5.26_mean = mean(temps5.26, na.rm = TRUE),
      .groups = 'drop'
    )
  
  # Cumulative values
  avgTbl <- avgTbl %>%
    mutate(
      cuml0.26 = cumsum(temps0.26_mean),
      cuml5.26 = cumsum(temps5.26_mean)
    )
  
  return(avgTbl)
}

```

Calculation of temperature and GDD, using the location and date as reference 
```{r pressure, echo=FALSE}
library(openmeteo)
library(purrr)  
library(lubridate)

dates_Neo22 <- seq(from = as.Date("2022-01-20"), to = as.Date("2022-07-16"), by = "day")
Neo22_temp<-thermTimeprofile(loc=c(39.23248,-88.37806),  plntDate="2022-01-20", date="2022-07-16")
Neo22_temp$gdd<-NULL
Neo22_temp$gdd<-gdd_calculator(loc=c(39.23248,-88.37806),  plntDate="2022-01-20", date=dates_Neo22)

dates_Neo23 <- seq(from = as.Date("2023-01-08"), to = as.Date("2023-07-16"), by = "day")
Neo23_temp<-thermTimeprofile(loc=c(39.23248,-88.37806),  plntDate="2023-01-08", date="2023-07-16")
Neo23_temp$gdd<-NULL
Neo23_temp$gdd<-gdd_calculator(loc=c(39.23248,-88.37806),  plntDate="2023-01-08", date=dates_Neo23)



dates_Urb22 <- seq(from = as.Date("2021-12-27"), to = as.Date("2022-07-16"), by = "day")
Urb22_temp<-thermTimeprofile(loc=c(39.31948,-88.45283),  plntDate="2021-12-27", date="2022-07-16")
Urb22_temp$gdd<-NULL
Urb22_temp$gdd<-gdd_calculator(loc=c(39.31948,-88.45283),  plntDate="2021-12-27", date=dates_Urb22)

dates_Urb23 <- seq(from = as.Date("2022-12-29"), to = as.Date("2023-07-16"), by = "day")
Urb23_temp<-thermTimeprofile(loc=c(39.31948,-88.45283),  plntDate="2022-12-29", date="2023-07-16")
Urb23_temp$gdd<-NULL
Urb23_temp$gdd<-gdd_calculator(loc=c(39.31948,-88.45283),  plntDate="2022-12-29", date=dates_Urb23)


dates_StP22 <- seq(from = as.Date("2021-12-28"), to = as.Date("2022-07-16"), by = "day")
StP22_temp<-thermTimeprofile(loc=c(38.87208, -88.88265),  plntDate="2021-12-28", date="2022-07-16")
StP22_temp$gdd<-NULL
StP22_temp$gdd<-gdd_calculator(loc=c(38.87208, -88.88265),  plntDate="2021-12-28", date=dates_StP22)


dates_StP23 <- seq(from = as.Date("2023-01-16"), to = as.Date("2023-07-16"), by = "day")
StP23_temp<-thermTimeprofile(loc=c(38.90354, -88.87827),  plntDate="2023-01-16", date="2023-07-16")
StP23_temp$gdd<-NULL
StP23_temp$gdd<-gdd_calculator(loc=c(38.90354, -88.87827),  plntDate="2023-01-16", date=dates_StP23)


```




```{r}
# Define the Julian values to keep
julian_Neo22 <- c(104, 117, 124, 130, 132, 137, 151, 164, 165)

# Filter the data
Neo22_flights <- Neo22_temp %>%
  filter(julian %in% julian_Neo22)

# Define the Julian values to keep
julian_Neo23 <- c(52, 96, 103,  114, 123, 129, 137, 152, 156, 164)

# Filter the data
Neo23_flights <- Neo23_temp %>%
  filter(julian %in% julian_Neo23)

# Define the Julian values to keep
julian_Urb22 <- c(106, 122, 127,  133, 139, 143, 148, 154, 160, 166)

# Filter the data
Urb22_flights <- Urb22_temp %>%
  filter(julian %in% julian_Urb22)


# Define the Julian values to keep
julian_Urb23 <- c(61, 97, 102, 116, 124, 130, 138, 150, 159, 165)

# Filter the data
Urb23_flights <- Urb23_temp %>%
  filter(julian %in% julian_Urb23)


# Define the Julian values to keep
julian_StP22 <- c(104, 117, 124, 130, 132, 137, 151, 165)

# Filter the data
StP22_flights <- StP22_temp %>%
  filter(julian %in% julian_StP22)


# Define the Julian values to keep
julian_StP23 <- c(52,74,96,103,114, 123, 129, 137, 152, 156, 164)

# Filter the data
StP23_flights <- StP23_temp %>%
  filter(julian %in% julian_StP23)

```

Create a new dataframe with the columns to see the exchange between min-max gdd, the difference between the last flight and the first one
```{r}
Dif_dat_gdd <- data.frame(
  location = c("Neoga", "Neoga", "Urbana", "Urbana", "St. Peter", "St. Peter"),
   year = c(2022, 2023, 2022, 2023, 2022, 2023),
  samples = c(9, 10, 10, 10, 8, 11),
  initial_date = as.Date(c("2022-04-14", "2023-02-21", "2022-04-16", 
                          "2023-03-02", "2022-04-14", "2023-02-21")),
  final_date = as.Date(c("2022-06-14", "2023-06-13", "2022-06-15", 
                        "2023-06-14", "2022-06-14", "2023-06-13")),
  min_gdd = c(421.05, 153.85, 492.80, 283.90, 523.35, 146.50),
  max_gdd = c(1542.10, 1634.10, 1616.75, 1722.80, 1689.70, 1691.75)
)

# Add columns of differences
Dif_dat_gdd <- Dif_dat_gdd %>%
  mutate(
    gdd_range = max_gdd - min_gdd,
    duration_days = as.numeric(final_date - initial_date),
    experiment = paste(location, year, sep = "_")
  )

```

Statistics by year
```{r Statistics by year}
stats_by_year <- Dif_dat_gdd %>%
  group_by(year) %>%
  summarise(
    n = n(),
    min_gdd_mean = mean(min_gdd),
    min_gdd_sd = sd(min_gdd),
    max_gdd_mean = mean(max_gdd),
    max_gdd_sd = sd(max_gdd),
    range_mean = mean(gdd_range),
    range_sd = sd(gdd_range),
    .groups = 'drop'
  )
```

Statistics by location
```{r Statistics by location} 
stats_by_location <-  Dif_dat_gdd %>%
  group_by(location) %>%
  summarise(
    across(c(min_gdd, max_gdd, gdd_range), 
           list(mean = mean, sd = sd), 
           .names = "{.col}_{.fn}"),
    .groups = 'drop'
  )
```

Check if the years are significant different in terms of gdd 
```{r}
data_2022 <- Dif_dat_gdd %>% filter(year == 2022)
data_2023 <- Dif_dat_gdd %>% filter(year == 2023)

# Test t para GDD mínimo
t.test(data_2022$min_gdd, data_2023$min_gdd, 
                     var.equal = FALSE, alternative = "two.sided")

# Test t para GDD máximo
t.test(data_2022$max_gdd, data_2023$max_gdd, 
                     var.equal = FALSE, alternative = "two.sided")

# Test t para rango de GDD
t.test(data_2022$gdd_range, data_2023$gdd_range, 
                       var.equal = FALSE, alternative = "two.sided")
```
```{r}
Dif_dat_gdd %>%
  group_by(year) %>%
  summarise(
    min_gdd_mean = mean(min_gdd),
    min_gdd_se = sd(min_gdd) / sqrt(n()),
    max_gdd_mean = mean(max_gdd),
    max_gdd_se = sd(max_gdd) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  pivot_longer(cols = contains("_mean"), 
               names_to = "variable", 
               values_to = "mean") %>%
  mutate(
    se = ifelse(grepl("min", variable), 
                rep(c(stats_by_year$min_gdd_sd[1]/sqrt(3), 
                     stats_by_year$min_gdd_sd[2]/sqrt(3))), 
                rep(c(stats_by_year$max_gdd_sd[1]/sqrt(3), 
                     stats_by_year$max_gdd_sd[2]/sqrt(3)))),
    variable = case_when(
      variable == "min_gdd_mean" ~ "GDD Min.",
      variable == "max_gdd_mean" ~ "GDD Max."
    )
  ) %>%
  ggplot(aes(x = factor(year), y = mean, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                position = position_dodge(0.9), width = 0.25) +
  labs(title = "Comparison of GDD for Year",
       x = "Year", y = "GDD", fill = "Type of GDD") +
  theme_minimal() +
  scale_fill_manual(values = c("GDD Min." = "#FF6B6B", "GDD Max." = "#4ECDC4"))
```

Calculate distance with the closest gdd neightbor
```{r}
library(dplyr)
library(tidyr)
# Redefine the function with explicit dplyr references
find_closest_gdd <- function(data1, value) {
  # Calculate the absolute distance and return the closest row
  closest_row <- data1 %>%
    dplyr::mutate(distance = abs(gdd - value)) %>%
    dplyr::arrange(distance) %>%
    dplyr::slice(1)  # Get the row with the smallest distance
  
  return(as.list(closest_row))  # Return as a list for rowwise compatibility
}
```


Find the closest neightboor in each location comparison  with St. Peter 2023
```{r  }
library(tidyverse)
filtered_results <- StP23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo22_temp, gdd))) %>%
  ungroup()

# Convert the list column to a dataframe
StP23.Neo22 <- do.call(rbind, filtered_results$closest_row)
  
####Neo23
filtered_results <- StP23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo23_temp, gdd))) %>%
  ungroup()

# Convert the list column to a dataframe
StP23.Neo23 <- do.call(rbind, filtered_results$closest_row)

####Urb 22
filtered_results <- StP23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb22_temp, gdd))) %>%
  ungroup()

# Convert the list column to a dataframe
StP23.Urb22 <- do.call(rbind, filtered_results$closest_row)

####Urb 23
filtered_results <- StP23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb23_temp, gdd))) %>%
  ungroup()

# Convert the list column to a dataframe
StP23.Urb23 <- do.call(rbind, filtered_results$closest_row)


####StP 22
filtered_results <- StP23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP22_temp, gdd))) %>%
  ungroup()
# Convert the list column to a dataframe
StP23.StP22 <- do.call(rbind, filtered_results$closest_row)

```


```{r St. Peter 2022}
####Neo22
filtered_results <- StP22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo22_temp, gdd))) %>%
  ungroup()

# Convert the list column to a dataframe
StP22.Neo22 <- do.call(rbind, filtered_results$closest_row)

####Neo23
filtered_results <- StP22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo23_temp, gdd)))

# Convert the list column to a dataframe
StP22.Neo23 <- do.call(rbind, filtered_results$closest_row)

####Urb 22
filtered_results <- StP22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb22_temp, gdd)))

# Convert the list column to a dataframe
StP22.Urb22 <- do.call(rbind, filtered_results$closest_row)

####Urb 23
filtered_results <- StP22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb23_temp, gdd)))

# Convert the list column to a dataframe
StP22.Urb23 <- do.call(rbind, filtered_results$closest_row)


####StP 22
filtered_results <- StP22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP23_temp, gdd)))

# Convert the list column to a dataframe
StP22.StP23 <- do.call(rbind, filtered_results$closest_row)

```


```{r Urbana 2022}
####Neo22
filtered_results <- Urb22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo22_temp, gdd)))

# Convert the list column to a dataframe
Urb22.Neo22 <- do.call(rbind, filtered_results$closest_row)

####Neo23
filtered_results <- Urb22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo23_temp, gdd)))

# Convert the list column to a dataframe
Urb22.Neo23 <- do.call(rbind, filtered_results$closest_row)

####StP 22
filtered_results <- Urb22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP22_temp, gdd)))

# Convert the list column to a dataframe
Urb22.StP22 <- do.call(rbind, filtered_results$closest_row)

####Urb 23
filtered_results <- Urb22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb23_temp, gdd)))

# Convert the list column to a dataframe
Urb22.Urb23 <- do.call(rbind, filtered_results$closest_row)


####StP 22
filtered_results <- Urb22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP23_temp, gdd)))

# Convert the list column to a dataframe
Urb22.StP23 <- do.call(rbind, filtered_results$closest_row)

```

```{r Urbana 2023}
####Neo22
filtered_results <- Urb23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo22_temp, gdd)))

# Convert the list column to a dataframe
Urb23.Neo22 <- do.call(rbind, filtered_results$closest_row)

####Neo23
filtered_results <- Urb23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo23_temp, gdd)))

# Convert the list column to a dataframe
Urb23.Neo23 <- do.call(rbind, filtered_results$closest_row)

####StP 22
filtered_results <- Urb23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP22_temp, gdd)))

# Convert the list column to a dataframe
Urb23.StP22 <- do.call(rbind, filtered_results$closest_row)

####Urb 22
filtered_results <- Urb23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb22_temp, gdd)))

# Convert the list column to a dataframe
Urb23.Urb22 <- do.call(rbind, filtered_results$closest_row)


####StP 22
filtered_results <- Urb23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP23_temp, gdd)))

# Convert the list column to a dataframe
Urb23.StP23 <- do.call(rbind, filtered_results$closest_row)

```


```{r Neoga 2022}
####Urb23
filtered_results <- Neo22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb23_temp, gdd)))

# Convert the list column to a dataframe
Neo22.Urb23 <- do.call(rbind, filtered_results$closest_row)

####Neo23
filtered_results <- Neo22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo23_temp, gdd)))

# Convert the list column to a dataframe
Neo22.Neo23 <- do.call(rbind, filtered_results$closest_row)

####StP 22
filtered_results <- Neo22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP22_temp, gdd)))

# Convert the list column to a dataframe
Neo22.StP22 <- do.call(rbind, filtered_results$closest_row)

####Urb 22
filtered_results <- Neo22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb22_temp, gdd)))

# Convert the list column to a dataframe
Neo22.Urb22 <- do.call(rbind, filtered_results$closest_row)


####StP 23
filtered_results <- Neo22_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP23_temp, gdd)))

# Convert the list column to a dataframe
Neo22.StP23 <- do.call(rbind, filtered_results$closest_row)

```

```{r Neoga 2023}
####Urb23
filtered_results <- Neo23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb23_temp, gdd)))

# Convert the list column to a dataframe
Neo23.Urb23 <- do.call(rbind, filtered_results$closest_row)

####Neo22
filtered_results <- Neo23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Neo22_temp, gdd)))

# Convert the list column to a dataframe
Neo23.Neo22 <- do.call(rbind, filtered_results$closest_row)

####StP 22
filtered_results <- Neo23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP22_temp, gdd)))

# Convert the list column to a dataframe
Neo23.StP22 <- do.call(rbind, filtered_results$closest_row)

####Urb 22
filtered_results <- Neo23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(Urb22_temp, gdd)))

# Convert the list column to a dataframe
Neo23.Urb22 <- do.call(rbind, filtered_results$closest_row)


####StP 23
filtered_results <- Neo23_flights %>%
  rowwise() %>%
  mutate(closest_row = list(find_closest_gdd(StP23_temp, gdd)))

# Convert the list column to a dataframe
Neo23.StP23 <- do.call(rbind, filtered_results$closest_row)


```



Now we are going to alingn the dates according with gdd, that are closest with the observed data in each location 
```{r Every flight}
Neo22_flights$date<- as.numeric(Neo22_flights$date)

Neo22_allflights<-rbind(Neo22_flights, Urb22.Neo22[,1:8], Urb23.Neo22[,1:8], Neo23.Neo22[,1:8], StP22.Neo22[,1:8], StP23.Neo22[,1:8])

Neo23_flights$date<- as.numeric(Neo23_flights$date)

Neo23_allflights<-rbind(Neo23_flights, Urb22.Neo23[,1:8], Urb23.Neo23[,1:8], Neo22.Neo23[,1:8], StP22.Neo23[,1:8], StP23.Neo23[,1:8])

Urb22_flights$date<- as.numeric(Urb22_flights$date)

Urb22_allflights<-rbind(Urb22_flights, Neo23.Urb22[,1:8], Urb23.Urb22[,1:8], Neo22.Urb22[,1:8], StP22.Urb22[,1:8], StP23.Urb22[,1:8])

Urb23_flights$date<- as.numeric(Urb23_flights$date)

Urb23_allflights<-rbind(Urb23_flights, Neo22.Urb23[,1:8], Urb22.Urb23[,1:8], Neo23.Urb23[,1:8], StP22.Urb23[,1:8], StP23.Urb23[,1:8])

StP22_flights$date<- as.numeric(StP22_flights$date)

StP22_allflights<-rbind(StP22_flights, Neo22.StP22[,1:8], Urb23.StP22[,1:8], Neo23.StP22[,1:8], Urb22.StP22[,1:8], StP23.StP22[,1:8])

StP23_flights$date<- as.numeric(StP23_flights$date)

StP23_allflights<-rbind(StP23_flights, Neo22.StP23[,1:8], Urb22.StP23[,1:8], Neo23.StP23[,1:8], StP22.StP23[,1:8], Urb23.StP23[,1:8])


```

Sort the data 
```{r}

Neo22_allflights$gdd<-as.numeric(Neo22_allflights$gdd)
Neo22_allflights <- Neo22_allflights %>%
   arrange(gdd)

Neo23_allflights$gdd<-as.numeric(Neo23_allflights$gdd)
Neo23_allflights <- Neo23_allflights %>%
   arrange(gdd)

Urb22_allflights$gdd<-as.numeric(Urb22_allflights$gdd)
Urb22_allflights <- Urb22_allflights %>%
   arrange(gdd)

Urb23_allflights$gdd<-as.numeric(Urb23_allflights$gdd)
Urb23_allflights <- Urb23_allflights %>%
   arrange(gdd)

StP22_allflights$gdd<-as.numeric(StP22_allflights$gdd)
StP22_allflights <- StP22_allflights %>%
   arrange(gdd)

StP23_allflights$gdd<-as.numeric(StP23_allflights$gdd)
StP23_allflights <- StP23_allflights %>%
   arrange(gdd)


```
