---
title: "Untitled"
author: "Christian Escoto"
date: "2025-07-01"
output: pdf_document
---


```{r}
library(dplyr)
library(asreml)
```


Load the data
```{r Data}
All<-read.csv("All.csv")
All$X<-NULL

All <- All %>%
  select(-c(starts_with("julian", ignore.case = FALSE)))

All <- All %>%
  select(-c(starts_with("temps5", ignore.case = FALSE)))

All <- All %>%
  select(-c(starts_with("cuml", ignore.case = FALSE)))

```



```{r cars}
N22<-All%>%
  filter(Location=="Neoga 2022")

N23<-All%>%
  filter(Location=="Neoga 2023")
# Replace infinite values with NA
N23[sapply(N23, is.infinite)] <- NA

N23<-na.omit(N23)

U22<-All%>%
  filter(Location=="Urbana 2022")
U22[sapply(U22, is.infinite)] <- NA

U22<-na.omit(U22)

U23<-All%>%
  filter(Location=="Urbana 2023")
U23[sapply(U23, is.infinite)] <- NA

U23<-na.omit(U23)

U23[sapply(U23, is.nan)] <- NA

U23<-na.omit(U23)

S22<-All%>%
  filter(Location=="St. Peter 2022")

S23<-All%>%
  filter(Location=="St. Peter 2023")
```

Now we are going to get a 80% percent 
```{r N23}
# Calculate the sample size (80% of the filtered rows)
sample1_N23 <- ceiling(0.2 * nrow(N23))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_N23<- N23[sample(nrow(N23), sample1_N23), ]

# Remove the sampled rows from the original data frame
sampled2_N23_copy <- N23[!rownames(N23) %in% rownames(sampled1_N23), ]
# Remove the sampled rows from the original data frame
sampled2_N23 <- N23[!rownames(N23) %in% rownames(sampled1_N23), ]

#Eliminate the unnecesary 
sampled2_N23$Location<-NULL
sampled2_N23$Unit<-NULL
sampled1_N23$Location<-NULL
sampled1_N23$Unit<-NULL


#Prepare the data to be used as the test
sampled1_N23.1<-sampled1_N23
sampled1_N23.1$Grain.yield<-NULL

sampled2_N23<-sapply(sampled2_N23, as.numeric)
sampled1_N23.1<-sapply(sampled1_N23.1, as.numeric)

sampled2_N23[is.infinite(sampled2_N23)]<-0
sampled2_N23[is.nan(sampled2_N23)]<-0
sampled2_N23[is.na(sampled2_N23)]<-0

sampled1_N23.1<-as.data.frame(sampled1_N23.1)
sampled2_N23<-as.data.frame(sampled2_N23)

```

```{r N22}
# Calculate the sample size (80% of the filtered rows)
sample1_N22 <- ceiling(0.2 * nrow(N22))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_N22<- N22[sample(nrow(N22), sample1_N22), ]

sampled2_N22_copy <- N22[!rownames(N22) %in% rownames(sampled1_N22), ]
# Remove the sampled rows from the original data frame
sampled2_N22 <- N22[!rownames(N22) %in% rownames(sampled1_N22), ]

#Eliminate the unnecesary 
sampled2_N22$Location<-NULL
sampled2_N22$Unit<-NULL
sampled1_N22$Location<-NULL
sampled1_N22$Unit<-NULL


#Prepare the data to be used as the test
sampled1_N22.1<-sampled1_N22
sampled1_N22.1$Grain.yield<-NULL

sampled2_N22<-sapply(sampled2_N22, as.numeric)
sampled1_N22.1<-sapply(sampled1_N22.1, as.numeric)

sampled2_N22[is.infinite(sampled2_N22)]<-0
sampled2_N22[is.nan(sampled2_N22)]<-0
sampled2_N22[is.na(sampled2_N22)]<-0

sampled1_N22.1<-as.data.frame(sampled1_N22.1)
sampled2_N22<-as.data.frame(sampled2_N22)

```

```{r U23}
# Calculate the sample size (80% of the filtered rows)
sample1_U23 <- ceiling(0.2 * nrow(U23))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_U23<- U23[sample(nrow(U23), sample1_U23), ]

sampled2_U23_copy <- U23[!rownames(U23) %in% rownames(sampled1_U23), ]
# Remove the sampled rows from the original data frame
sampled2_U23 <- U23[!rownames(U23) %in% rownames(sampled1_U23), ]

#Eliminate the unnecesary 
sampled2_U23$Location<-NULL
sampled2_U23$Unit<-NULL
sampled1_U23$Location<-NULL
sampled1_U23$Unit<-NULL


#Prepare the data to be used as the test
sampled1_U23.1<-sampled1_U23
sampled1_U23.1$Grain.yield<-NULL

sampled2_U23<-sapply(sampled2_U23, as.numeric)
sampled1_U23.1<-sapply(sampled1_U23.1, as.numeric)

sampled2_U23[is.infinite(sampled2_U23)]<-0
sampled2_U23[is.nan(sampled2_U23)]<-0
sampled2_U23[is.na(sampled2_U23)]<-0

sampled1_U23.1<-as.data.frame(sampled1_U23.1)
sampled2_U23<-as.data.frame(sampled2_U23)

```

```{r U22}
# Calculate the sample size (80% of the filtered rows)
sample1_U22 <- ceiling(0.2 * nrow(U22))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_U22<- U22[sample(nrow(U22), sample1_U22), ]

sampled2_U22_copy <- U22[!rownames(U22) %in% rownames(sampled1_U22), ]
# Remove the sampled rows from the original data frame
sampled2_U22 <- U22[!rownames(U22) %in% rownames(sampled1_U22), ]

#Eliminate the unnecesary 
sampled2_U22$Location<-NULL
sampled2_U22$Unit<-NULL
sampled1_U22$Location<-NULL
sampled1_U22$Unit<-NULL


#Prepare the data to be used as the test
sampled1_U22.1<-sampled1_U22
sampled1_U22.1$Grain.yield<-NULL

sampled2_U22<-sapply(sampled2_U22, as.numeric)
sampled1_U22.1<-sapply(sampled1_U22.1, as.numeric)

sampled2_U22[is.infinite(sampled2_U22)]<-0
sampled2_U22[is.nan(sampled2_U22)]<-0
sampled2_U22[is.na(sampled2_U22)]<-0

sampled1_U22.1<-as.data.frame(sampled1_U22.1)
sampled2_U22<-as.data.frame(sampled2_U22)

```

```{r S23}
# Calculate the sample size (80% of the filtered rows)
sample1_S23 <- ceiling(0.2 * nrow(S23))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_S23<- S23[sample(nrow(S23), sample1_S23), ]

sampled2_S23_copy <- S23[!rownames(S23) %in% rownames(sampled1_S23), ]
# Remove the sampled rows from the original data frame
sampled2_S23 <- S23[!rownames(S23) %in% rownames(sampled1_S23), ]

#Eliminate the unnecesary 
sampled2_S23$Location<-NULL
sampled2_S23$Unit<-NULL
sampled1_S23$Location<-NULL
sampled1_S23$Unit<-NULL


#Prepare the data to be used as the test
sampled1_S23.1<-sampled1_S23
sampled1_S23.1$Grain.yield<-NULL

sampled2_S23<-sapply(sampled2_S23, as.numeric)
sampled1_S23.1<-sapply(sampled1_S23.1, as.numeric)

sampled2_S23[is.infinite(sampled2_S23)]<-0
sampled2_S23[is.nan(sampled2_S23)]<-0
sampled2_S23[is.na(sampled2_S23)]<-0

sampled1_S23.1<-as.data.frame(sampled1_S23.1)
sampled2_S23<-as.data.frame(sampled2_S23)

```

```{r S22}
# Calculate the sample size (80% of the filtered rows)
sample1_S22 <- ceiling(0.2 * nrow(S22))

# Take a random sample of 80%
set.seed(123)  # Set seed for reproducibility
sampled1_S22<- S22[sample(nrow(S22), sample1_S22), ]


sampled2_S22_copy <- S22[!rownames(S22) %in% rownames(sampled1_S22), ]
# Remove the sampled rows from the original data frame
sampled2_S22 <- S22[!rownames(S22) %in% rownames(sampled1_S22), ]

#Eliminate the unnecesary 
sampled2_S22$Location<-NULL
sampled2_S22$Unit<-NULL
sampled1_S22$Location<-NULL
sampled1_S22$Unit<-NULL


#Prepare the data to be used as the test
sampled1_S22.1<-sampled1_S22
sampled1_S22.1$Grain.yield<-NULL

sampled2_S22<-sapply(sampled2_S22, as.numeric)
sampled1_S22.1<-sapply(sampled1_S22.1, as.numeric)

sampled2_S22[is.infinite(sampled2_S22)]<-0
sampled2_S22[is.nan(sampled2_S22)]<-0
sampled2_S22[is.na(sampled2_S22)]<-0

sampled1_S22.1<-as.data.frame(sampled1_S22.1)
sampled2_S22<-as.data.frame(sampled2_S22)

```

Predictions by each model
```{r catboost}
#Making the prediction
result_N23_cat <-cat_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", seed = 123)
result_N23_cat
result_N23_cat$RMSE
#472.87
result_N23_cat$correlation
#0.51


#Making the prediction
result_N22_cat <-cat_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", seed = 123)
result_N22_cat$RMSE
# 410.41
result_N22_cat$correlation
# 0.7


#Making the prediction
result_U23_cat <-cat_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", seed = 123)
result_U23_cat$RMSE
#  625.86
result_U23_cat$correlation
# 0.72

#Making the prediction
result_U22_cat <-cat_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", seed = 123)
result_U22_cat$RMSE
# 481.67
result_U22_cat$correlation
# 0.67

#Making the prediction
result_S23_cat <-cat_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", seed = 123)
result_S23_cat$RMSE
#   418.37
result_S23_cat$correlation
# 0.80

#Making the prediction
result_S22_cat <-cat_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", seed = 123)
result_S22_cat$RMSE
# 429.07
result_S22_cat$correlation
# 0.52


```

```{r mboost}

#Making the prediction
result_N23_mboost <-mboost_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", seed = 123)
result_N23_mboost$RMSE
#461.3
result_N23_mboost$correlation
#0.54


#Making the prediction
result_N22_mboost <-mboost_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", seed = 123)
result_N22_mboost$RMSE
# 390
result_N22_mboost$correlation
# 0.74


#Making the prediction
result_U23_mboost <-mboost_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", seed = 123)
result_U23_mboost$RMSE
#  608.6
result_U23_mboost$correlation
# 0.74

#Making the prediction
result_U22_mboost <-mboost_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", seed = 123)
result_U22_mboost$RMSE
# 463.16
result_U22_mboost$correlation
# 0.70

#Making the prediction
result_S23_mboost <-mboost_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", seed = 123)
result_S23_mboost$RMSE
# 384.41
result_S23_mboost$correlation
# 0.84

#Making the prediction
result_S22_mboost <-mboost_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", seed = 123)
result_S22_mboost$RMSE
# 404.17
result_S22_mboost$correlation
# 0.63


```

```{r pcr}
#Making the prediction
result_N23_pcr <-pcr_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", seed = 123)
result_N23_pcr$RMSE
#476.98
result_N23_pcr$correlation
#0.53

#Making the prediction
result_N22_pcr <-pcr_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", seed = 123)
result_N22_pcr$RMSE
# 402.34
result_N22_pcr$correlation
# 0.72

#Making the prediction
result_U23_pcr <-pcr_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", seed = 123)
result_U23_pcr$RMSE
#  622.21
result_U23_pcr$correlation
# 0.73

#Making the prediction
result_U22_pcr <-pcr_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", seed = 123)
result_U22_pcr$RMSE
# 453.88
result_U22_pcr$correlation
# 0.72

#Making the prediction
result_S23_pcr <-pcr_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", seed = 123)
result_S23_pcr$RMSE
# 390.86
result_S23_pcr$correlation
# 0.84

#Making the prediction
result_S22_pcr <-pcr_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", seed = 123)
result_S22_pcr$RMSE
# 421.75
result_S22_pcr$correlation
# 0.56
```

```{r xgb}
#Making the prediction
result_N23_xgb <-xgb_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", seed = 123)
result_N23_xgb$RMSE
#473.13
result_N23_xgb$correlation
#0.51


#Making the prediction
result_N22_xgb <-xgb_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", seed = 123)
result_N22_xgb$RMSE
# 411.73
result_N22_xgb$correlation
# 0.7


#Making the prediction
result_U23_xgb <-xgb_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", seed = 123)
result_U23_xgb$RMSE
#  608.93
result_U23_xgb$correlation
# 0.74

#Making the prediction
result_U22_xgb <-xgb_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", seed = 123)
result_U22_xgb$RMSE
# 485.22
result_U22_xgb$correlation
# 0.68

#Making the prediction
result_S23_xgb <-xgb_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", seed = 123)
result_S23_xgb$RMSE
# 424.91
result_S23_xgb$correlation
# 0.8

#Making the prediction
result_S22_xgb <-xgb_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", seed = 123)
result_S22_xgb$RMSE
# 427.91
result_S22_xgb$correlation
# 0.53
```

```{r rf}
#Making the prediction
result_N23_rf <-rf_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", seed = 123)
result_N23_rf$RMSE
#465.91
result_N23_rf$correlation
#0.53


#Making the prediction
result_N22_rf <-rf_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", seed = 123)
result_N22_rf$RMSE
# 406.75
result_N22_rf$correlation
# 0.7


#Making the prediction
result_U23_rf <-rf_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", seed = 123)
result_U23_rf$RMSE
#  629.45
result_U23_rf$correlation
# 0.73

#Making the prediction
result_U22_rf <-rf_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", seed = 123)
result_U22_rf$RMSE
# 478.04
result_U22_rf$correlation
# 0.68

#Making the prediction
result_S23_rf <-rf_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", seed = 123)
result_S23_rf$RMSE
# 414.39
result_S23_rf$correlation
# 0.8

#Making the prediction
result_S22_rf <-rf_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", seed = 123)
result_S22_rf$RMSE
# 434.95
result_S22_rf$correlation
# 0.53
```

```{r bayesA}
#Making the prediction
result_N23_bayesA <-bayesA_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield")
result_N23_bayesA$RMSE
#465.51
result_N23_bayesA$correlation
#0.54

#Making the prediction
result_N22_bayesA <-bayesA_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield")
result_N22_bayesA$RMSE
# 386.48
result_N22_bayesA$correlation
# 0.74

#Making the prediction
result_U23_bayesA <-bayesA_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield")
result_U23_bayesA$RMSE
#  688.6
result_U23_bayesA$correlation
# 0.67

#Making the prediction
result_U22_bayesA <-bayesA_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield")
result_U22_bayesA$RMSE
# 451.40
result_U22_bayesA$correlation
# 0.72

#Making the prediction
result_S23_bayesA <-bayesA_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield")
result_S23_bayesA$RMSE
# 375.94
result_S23_bayesA$correlation
# 0.85

#Making the prediction
result_S22_bayesA <-bayesA_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield")
result_S22_bayesA$RMSE
# 411.43
result_S22_bayesA$correlation
# 06053
```

```{r bayesL}
#Making the prediction
result_N23_bayesL <-bayesL_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield")
result_N23_bayesL$RMSE
#461.98
result_N23_bayesL$correlation
#0.55

#Making the prediction
result_N22_bayesL <-bayesL_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield")
result_N22_bayesL$RMSE
# 387.55
result_N22_bayesL$correlation
# 0.74

#Making the prediction
result_U23_bayesL <-bayesL_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield")
result_U23_bayesL$RMSE
#  683.33
result_U23_bayesL$correlation
# 0.68

#Making the prediction
result_U22_bayesL <-bayesL_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield")
result_U22_bayesL$RMSE
# 452.18
result_U22_bayesL$correlation
# 0.72

#Making the prediction
result_S23_bayesL <-bayesL_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield")
result_S23_bayesL$RMSE
# 377.32
result_S23_bayesL$correlation
# 0.85

#Making the prediction
result_S22_bayesL <-bayesL_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield")
result_S22_bayesL$RMSE
# 411.41
result_S22_bayesL$correlation
# 06
```

```{r bayesR}
#Making the prediction
result_N23_bayesR <-bayesR_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield")
result_N23_bayesR$RMSE
#465.44
result_N23_bayesR$correlation
#0.54

#Making the prediction
result_N22_bayesR <-bayesR_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield")
result_N22_bayesR$RMSE
# 384.9
result_N22_bayesR$correlation
# 0.74

#Making the prediction
result_U23_bayesR <-bayesR_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield")
result_U23_bayesR$RMSE
#  685.75
result_U23_bayesR$correlation
# 0.67

#Making the prediction
result_U22_bayesR <-bayesR_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield")
result_U22_bayesR$RMSE
# 451.16
result_U22_bayesR$correlation
# 0.72

#Making the prediction
result_S23_bayesR <-bayesR_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield")
result_S23_bayesR$RMSE
# 372.43
result_S23_bayesR$correlation
# 0.84

#Making the prediction
result_S22_bayesR <-bayesR_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield")
result_S22_bayesR$RMSE
# 411.19
result_S22_bayesR$correlation
# 0.6
```

```{r ll}
#Making the prediction
result_N23_ll <-lineal_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", alpha_value = 1)
result_N23_ll$RMSE
#467.23
result_N23_ll$correlation
#0.53

#Making the prediction
result_N22_ll <-lineal_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", alpha_value = 1)
result_N22_ll$RMSE
# 391.54
result_N22_ll$correlation
# 0.73

#Making the prediction
result_U23_ll <-lineal_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", alpha_value = 1)
result_U23_ll$RMSE
#  648.2
result_U23_ll$correlation
# 0.7

#Making the prediction
result_U22_ll <-lineal_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", alpha_value = 1)
result_U22_ll$RMSE
# 451.95
result_U22_ll$correlation
# 0.72

#Making the prediction
result_S23_ll <-lineal_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", alpha_value = 1)
result_S23_ll$RMSE
# 380.64
result_S23_ll$correlation
# 0.84

#Making the prediction
result_S22_ll <-lineal_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", alpha_value = 1)
result_S22_ll$RMSE
# 419.04
result_S22_ll$correlation
# 0.58
```

```{r lr}
#Making the prediction
result_N23_lr <-lineal_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", alpha_value = 1)
result_N23_lr$RMSE
#467.23
result_N23_lr$correlation
#0.53

#Making the prediction
result_N22_lr <-lineal_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", alpha_value = 1)
result_N22_lr$RMSE
# 391.54
result_N22_lr$correlation
# 0.73

#Making the prediction
result_U23_lr <-lineal_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", alpha_value = 1)
result_U23_lr$RMSE
#  648.2
result_U23_lr$correlation
# 0.7

#Making the prediction
result_U22_lr <-lineal_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", alpha_value = 1)
result_U22_lr$RMSE
# 451.95
result_U22_lr$correlation
# 0.72

#Making the prediction
result_S23_lr <-lineal_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", alpha_value = 1)
result_S23_lr$RMSE
# 380.64
result_S23_lr$correlation
# 0.84

#Making the prediction
result_S22_lr <-lineal_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", alpha_value = 1)
result_S22_lr$RMSE
# 419.04
result_S22_lr$correlation
# 0.58
```

```{r le}
#Making the prediction
result_N23_le <-lineal_model(train_data = sampled1_N23, test_data = sampled2_N23,  target_variable= "Grain.yield", alpha_value = 0.5)
result_N23_le$RMSE
#467.17
result_N23_le$correlation
#0.53

#Making the prediction
result_N22_le <-lineal_model(train_data = sampled1_N22, test_data = sampled2_N22,  target_variable= "Grain.yield", alpha_value = 0.5)
result_N22_le$RMSE
# 390.72
result_N22_le$correlation
# 0.74

#Making the prediction
result_U23_le <-lineal_model(train_data = sampled1_U23, test_data = sampled2_U23,  target_variable= "Grain.yield", alpha_value = 0.5)
result_U23_le$RMSE
#  669.55
result_U23_le$correlation
# 0.68

#Making the prediction
result_U22_le <-lineal_model(train_data = sampled1_U22, test_data = sampled2_U22,  target_variable= "Grain.yield", alpha_value = 0.5)
result_U22_le$RMSE
# 451.96
result_U22_le$correlation
# 0.72

#Making the prediction
result_S23_le <-lineal_model(train_data = sampled1_S23, test_data = sampled2_S23,  target_variable= "Grain.yield", alpha_value = 0.5)
result_S23_le$RMSE
# 380.40
result_S23_le$correlation
# 0.84

#Making the prediction
result_S22_le <-lineal_model(train_data = sampled1_S22, test_data = sampled2_S22,  target_variable= "Grain.yield", alpha_value = 0.5)
result_S22_le$RMSE
# 418.33
result_S22_le$correlation
# 0.58
```

Checking the predictions' results in each location
```{r N22}
models.N22<- c('result_N22_cat', 'result_N22_mboost', 'result_N22_rf', 'result_N22_xgb', 'result_N22_pcr', 'result_N22_bayesA', 'result_N22_bayesL', 'result_N22_bayesR', 'result_N22_ll', 'result_N22_le', 'result_N22_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.N22)){
  pa_vec<- append(pa_vec, get(models.N22[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.N22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.N22)){
  pa_vec<- append(pa_vec, get(models.N22[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.N22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_N22_cat$predictions)

summary(result_N22_xgb$predictions)

summary(result_N22_mboost$predictions)

summary(result_N22_rf$predictions)

summary(result_N22_pcr$predictions)

summary(result_N22_bayesA$predictions)

summary(result_N22_bayesL$predictions)

summary(result_N22_le$predictions)

summary(result_N22_lr$predictions)

summary(result_N22_ll$predictions)
```

```{r N23}
models.N23<- c('result_N23_cat', 'result_N23_mboost', 'result_N23_rf', 'result_N23_xgb', 'result_N23_pcr', 'result_N23_bayesA', 'result_N23_bayesL', 'result_N23_bayesR', 'result_N23_ll', 'result_N23_le', 'result_N23_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.N23)){
  pa_vec<- append(pa_vec, get(models.N23[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.N23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.N23)){
  pa_vec<- append(pa_vec, get(models.N23[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.N23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_N23_cat$predictions)

summary(result_N23_xgb$predictions)

summary(result_N23_mboost$predictions)

summary(result_N23_rf$predictions)

summary(result_N23_pcr$predictions)

summary(result_N23_bayesA$predictions)

summary(result_N23_bayesL$predictions)

summary(result_N23_le$predictions)

summary(result_N23_lr$predictions)

summary(result_N23_ll$predictions)
```

```{r U22}
models.U22<- c('result_U22_cat', 'result_U22_mboost', 'result_U22_rf', 'result_U22_xgb', 'result_U22_pcr', 'result_U22_bayesA', 'result_U22_bayesL', 'result_U22_bayesR', 'result_U22_ll', 'result_U22_le', 'result_U22_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.U22)){
  pa_vec<- append(pa_vec, get(models.U22[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.U22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.U22)){
  pa_vec<- append(pa_vec, get(models.U22[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.U22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_U22_cat$predictions)

summary(result_U22_xgb$predictions)

summary(result_U22_mboost$predictions)

summary(result_U22_rf$predictions)

summary(result_U22_pcr$predictions)

summary(result_U22_bayesA$predictions)

summary(result_U22_bayesL$predictions)

summary(result_U22_bayesR$predictions)

summary(result_U22_le$predictions)

summary(result_U22_lr$predictions)

summary(result_U22_ll$predictions)
```

```{r U23}
models.U23<- c('result_U23_cat', 'result_U23_mboost', 'result_U23_rf', 'result_U23_xgb', 'result_U23_pcr', 'result_U23_bayesA', 'result_U23_bayesL', 'result_U23_bayesR', 'result_U23_ll', 'result_U23_le', 'result_U23_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.U23)){
  pa_vec<- append(pa_vec, get(models.U23[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.U23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.U23)){
  pa_vec<- append(pa_vec, get(models.U23[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.U23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_U23_cat$predictions)

summary(result_U23_xgb$predictions)

summary(result_U23_mboost$predictions)

summary(result_U23_rf$predictions)

summary(result_U23_pcr$predictions)

summary(result_U23_bayesA$predictions)

summary(result_U23_bayesL$predictions)

summary(result_U23_bayesR$predictions)

summary(result_U23_le$predictions)

summary(result_U23_lr$predictions)

summary(result_U23_ll$predictions)
```
```{r S22}
models.S22<- c('result_S22_cat', 'result_S22_mboost', 'result_S22_rf', 'result_S22_xgb', 'result_S22_pcr', 'result_S22_bayesA', 'result_S22_bayesL', 'result_S22_bayesR', 'result_S22_ll', 'result_S22_le', 'result_S22_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.S22)){
  pa_vec<- append(pa_vec, get(models.S22[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.S22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.S22)){
  pa_vec<- append(pa_vec, get(models.S22[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.S22, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_S22_cat$predictions)

summary(result_S22_xgb$predictions)

summary(result_S22_mboost$predictions)

summary(result_S22_rf$predictions)

summary(result_S22_pcr$predictions)

summary(result_S22_bayesA$predictions)

summary(result_S22_bayesL$predictions)

summary(result_S22_bayesR$predictions)

summary(result_S22_le$predictions)

summary(result_S22_lr$predictions)

summary(result_S22_ll$predictions)
```

```{r S23}
models.S23<- c('result_S23_cat', 'result_S23_mboost', 'result_S23_rf', 'result_S23_xgb', 'result_S23_pcr', 'result_S23_bayesA', 'result_S23_bayesL', 'result_S23_bayesR', 'result_S23_ll', 'result_S23_le', 'result_S23_lr')

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.S23)){
  pa_vec<- append(pa_vec, get(models.S23[i])$correlation)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.S23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

#loop to get results into a vector
pa_vec<-c()
for(i in 1:length(models.S23)){
  pa_vec<- append(pa_vec, get(models.S23[i])$RMSE)
}

#create data frame with model and predictive ability
Rtable<- data.frame(Model=models.S23, Pred_Ability= pa_vec)

#display table
knitr::kable(Rtable[order(Rtable$Pred_Ability, decreasing=TRUE),], digits=2, row.names=FALSE)

summary(result_S23_cat$predictions)

summary(result_S23_xgb$predictions)

summary(result_S23_mboost$predictions)

summary(result_S23_rf$predictions)

summary(result_S23_pcr$predictions)

summary(result_S23_bayesA$predictions)

summary(result_S23_bayesL$predictions)

summary(result_S23_bayesR$predictions)

summary(result_S23_le$predictions)

summary(result_S23_lr$predictions)

summary(result_S23_ll$predictions)
```

Upload the information of the plants in terms of blocks, units, etc.
```{r Plants info}
Plant_info<-read.csv("Plant.info.csv")
```


Include all predictions in the same dataframe
```{r N23}
sampled2_N23_copy$catboost<-result_N23_cat$predictions
sampled2_N23_copy$xgboost<-result_N23_xgb$predictions
sampled2_N23_copy$mboost<-result_N23_mboost$predictions
sampled2_N23_copy$rf<-result_N23_rf$predictions
sampled2_N23_copy$pcr<-result_N23_pcr$predictions
sampled2_N23_copy$bayesA<-result_N23_bayesA$predictions
sampled2_N23_copy$bayesL<-result_N23_bayesL$predictions
sampled2_N23_copy$bayesR<-result_N23_bayesR$predictions
sampled2_N23_copy$le<-result_N23_le$predictions
sampled2_N23_copy$lr<-result_N23_lr$predictions
sampled2_N23_copy$ll<-result_N23_ll$predictions
```


Merge with plant information
```{r}
N23.All<-merge(sampled2_N23_copy, Plant_info, by="Unit")
N23.All$Location.y<-NULL
N23.All<-rename(N23.All, geno = germplasmDbId)
N23.All<-rename(N23.All, Column=colNumber)
N23.All<-rename(N23.All, Row=rowNumber)
N23.All<-rename(N23.All, Rep=replicate)

N23.All$Column<-as.factor(N23.All$Column)
N23.All$Row<-as.factor(N23.All$Row)
N23.All$geno<-as.factor(N23.All$geno)
N23.All$blockNumber<-as.factor(N23.All$blockNumber)
N23.All$Rep<-as.factor(N23.All$Rep)
```


```{r N22}
sampled2_N22_copy$catboost<-result_N22_cat$predictions
sampled2_N22_copy$xgboost<-result_N22_xgb$predictions
sampled2_N22_copy$mboost<-result_N22_mboost$predictions
sampled2_N22_copy$rf<-result_N22_rf$predictions
sampled2_N22_copy$pcr<-result_N22_pcr$predictions
sampled2_N22_copy$bayesA<-result_N22_bayesA$predictions
sampled2_N22_copy$bayesL<-result_N22_bayesL$predictions
sampled2_N22_copy$bayesR<-result_N22_bayesR$predictions
sampled2_N22_copy$le<-result_N22_le$predictions
sampled2_N22_copy$lr<-result_N22_lr$predictions
sampled2_N22_copy$ll<-result_N22_ll$predictions
```

Merge with plant information
```{r}
N22.All<-merge(sampled2_N22_copy, Plant_info, by="Unit")
N22.All$Location.y<-NULL
N22.All<-rename(N22.All, geno = germplasmDbId)
N22.All<-rename(N22.All, Column=colNumber)
N22.All<-rename(N22.All, Row=rowNumber)
N22.All<-rename(N22.All, Rep=replicate)

N22.All$Column<-as.factor(N22.All$Column)
N22.All$Row<-as.factor(N22.All$Row)
N22.All$geno<-as.factor(N22.All$geno)
N22.All$blockNumber<-as.factor(N22.All$blockNumber)
N22.All$Rep<-as.factor(N22.All$Rep)

```

```{r U22}
sampled2_U22_copy$catboost<-result_U22_cat$predictions
sampled2_U22_copy$xgboost<-result_U22_xgb$predictions
sampled2_U22_copy$mboost<-result_U22_mboost$predictions
sampled2_U22_copy$rf<-result_U22_rf$predictions
sampled2_U22_copy$pcr<-result_U22_pcr$predictions
sampled2_U22_copy$bayesA<-result_U22_bayesA$predictions
sampled2_U22_copy$bayesL<-result_U22_bayesL$predictions
sampled2_U22_copy$bayesR<-result_U22_bayesR$predictions
sampled2_U22_copy$le<-result_U22_le$predictions
sampled2_U22_copy$lr<-result_U22_lr$predictions
sampled2_U22_copy$ll<-result_U22_ll$predictions
```

Merge with plant information
```{r}
U22.All<-merge(sampled2_U22_copy, Plant_info, by="Unit")
U22.All$Location.y<-NULL
U22.All<-rename(U22.All, geno = germplasmDbId)
U22.All<-rename(U22.All, Column=colNumber)
U22.All<-rename(U22.All, Row=rowNumber)
U22.All<-rename(U22.All, Rep=replicate)

U22.All$Column<-as.factor(U22.All$Column)
U22.All$Row<-as.factor(U22.All$Row)
U22.All$geno<-as.factor(U22.All$geno)
U22.All$blockNumber<-as.factor(U22.All$blockNumber)
U22.All$Rep<-as.factor(U22.All$Rep)


```

```{r U23}
sampled2_U23_copy$catboost<-result_U23_cat$predictions
sampled2_U23_copy$xgboost<-result_U23_xgb$predictions
sampled2_U23_copy$mboost<-result_U23_mboost$predictions
sampled2_U23_copy$rf<-result_U23_rf$predictions
sampled2_U23_copy$pcr<-result_U23_pcr$predictions
sampled2_U23_copy$bayesA<-result_U23_bayesA$predictions
sampled2_U23_copy$bayesL<-result_U23_bayesL$predictions
sampled2_U23_copy$bayesR<-result_U23_bayesR$predictions
sampled2_U23_copy$le<-result_U23_le$predictions
sampled2_U23_copy$lr<-result_U23_lr$predictions
sampled2_U23_copy$ll<-result_U23_ll$predictions
```

Merge with plant information
```{r}
U23.All<-merge(sampled2_U23_copy, Plant_info, by="Unit")
U23.All$Location.y<-NULL
U23.All<-rename(U23.All, geno = germplasmDbId)
U23.All<-rename(U23.All, Column=colNumber)
U23.All<-rename(U23.All, Row=rowNumber)
U23.All<-rename(U23.All, Rep=replicate)

U23.All$Column<-as.factor(U23.All$Column)
U23.All$Row<-as.factor(U23.All$Row)
U23.All$geno<-as.factor(U23.All$geno)
U23.All$blockNumber<-as.factor(U23.All$blockNumber)
U23.All$Rep<-as.factor(U23.All$Rep)
```


```{r S23}
sampled2_S23_copy$catboost<-result_S23_cat$predictions
sampled2_S23_copy$xgboost<-result_S23_xgb$predictions
sampled2_S23_copy$mboost<-result_S23_mboost$predictions
sampled2_S23_copy$rf<-result_S23_rf$predictions
sampled2_S23_copy$pcr<-result_S23_pcr$predictions
sampled2_S23_copy$bayesA<-result_S23_bayesA$predictions
sampled2_S23_copy$bayesL<-result_S23_bayesL$predictions
sampled2_S23_copy$bayesR<-result_S23_bayesR$predictions
sampled2_S23_copy$le<-result_S23_le$predictions
sampled2_S23_copy$lr<-result_S23_lr$predictions
sampled2_S23_copy$ll<-result_S23_ll$predictions
```

Merge with plant information
```{r}
S23.All<-merge(sampled2_S23_copy, Plant_info, by="Unit")
S23.All$Location.y<-NULL
S23.All<-rename(S23.All, geno = germplasmDbId)
S23.All<-rename(S23.All, Column=colNumber)
S23.All<-rename(S23.All, Row=rowNumber)
S23.All<-rename(S23.All, Rep=replicate)

S23.All$Column<-as.factor(S23.All$Column)
S23.All$Row<-as.factor(S23.All$Row)
S23.All$geno<-as.factor(S23.All$geno)
S23.All$blockNumber<-as.factor(S23.All$blockNumber)
S23.All$Rep<-as.factor(S23.All$Rep)
```


```{r S22}
sampled2_S22_copy$catboost<-result_S22_cat$predictions
sampled2_S22_copy$xgboost<-result_S22_xgb$predictions
sampled2_S22_copy$mboost<-result_S22_mboost$predictions
sampled2_S22_copy$rf<-result_S22_rf$predictions
sampled2_S22_copy$pcr<-result_S22_pcr$predictions
sampled2_S22_copy$bayesA<-result_S22_bayesA$predictions
sampled2_S22_copy$bayesL<-result_S22_bayesL$predictions
sampled2_S22_copy$bayesR<-result_S22_bayesR$predictions
sampled2_S22_copy$le<-result_S22_le$predictions
sampled2_S22_copy$lr<-result_S22_lr$predictions
sampled2_S22_copy$ll<-result_S22_ll$predictions
```

Merge with plant information
```{r}
S22.All<-merge(sampled2_S22_copy, Plant_info, by="Unit")
S22.All$Location.y<-NULL
S22.All<-rename(S22.All, geno = germplasmDbId)
S22.All<-rename(S22.All, Column=colNumber)
S22.All<-rename(S22.All, Row=rowNumber)
S22.All<-rename(S22.All, Rep=replicate)

S22.All$Column<-as.factor(S22.All$Column)
S22.All$Row<-as.factor(S22.All$Row)
S22.All$geno<-as.factor(S22.All$geno)
S22.All$blockNumber<-as.factor(S22.All$blockNumber)
S22.All$Rep<-as.factor(S22.All$Rep)
```
